---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
read.csv("activity.csv")
library(readr)
activity <- read_csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
totalsteps <-aggregate(steps ~ date, activity, FUN = sum)
```
#### Making a Histogram
```{r}
hist(totalsteps$steps, ylab = "Number of Days", xlab = "Number of steps", col = "red", main = "Total Steps Daily")
```
#### Finding the mean and median
```{r}
meansteps<-mean(totalsteps$steps)
mediansteps<-median(totalsteps$steps)
```
## What is the average daily activity pattern?
```{r}
library(ggplot2)
meanStepsInt<-aggregate(steps ~ interval, activity, mean)
ggplot(data = meanStepsInt, aes(x=interval, y=steps))+
  geom_line()+
  ggtitle("Average Activity Pattern")+
  xlab("5-minute Interval")+
  ylab("Average number of steps")
```
#### 5-Minute interval containing maximum steps
```{r}
MaxIntv<-meanStepsInt[which.max(meanStepsInt$steps),]
```

## Imputing missing values
Use the **mean interval** to fill the missing values 
```{r}
newactivity<-activity
for (i in meanStepsInt$interval) {
  newactivity[newactivity$interval == i & is.na(newactivity$steps), ]$steps <-
    meanStepsInt$steps[meanStepsInt$interval ==i]
}
head(newactivity)
sum(is.na(newactivity)) #####Making sure there are no missing values
```

#### Creating a Historgram for the new Data
```{r}
newactivitytotal<-aggregate(steps ~ date, data = newactivity, sum, na.rm=TRUE)
hist(newactivitytotal$steps, ylab = "Number of Days", xlab = "Number of steps", col="blue", main = "New Daily Steps")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
daytypes<-function(date) {
  days<-weekdays(date)
  if (days %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return("weekday")
  else if (days %in% c('Saturday', 'Sunday'))
      return("weekend")
  else
      stop("Invalid Date Format")
}
newactivity$date <-as.Date(newactivity$date)
newactivity$day <-sapply(newactivity$date, FUN = daytypes)

meanstepsdaily<- aggregate(steps ~ interval + day, newactivity, mean)
ggplot(data = meanstepsdaily, aes(x=interval, y=steps))+
  geom_line()+
  facet_grid(day ~ .)+
  ggtitle("Average Daily Activity Pattern")+
  xlab("5- Minute Interval")+
  ylab("Average Number of Steps")
```